True Lies

### 2018-11-05 GSV-comment-brief.Rmd
check differences between their results and my calculations; make GSV-heads correspond to what they say, not what they do; add bias table; add power calculations; update readme; stop using workaround for hdrcde::hdr (thanks @robjhyndman); look at benndorf et al. 2017

---
title: "True Lies:"
subtitle: "Comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{setspace}
  - \usepackage{placeins}
  - \onehalfspacing
  - \usepackage{fontspec}
output:
  pdf_document:
    md_extensions: +raw_attribute
    latex_engine: xelatex
  tufte::tufte_handout:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_handout2:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
fontsize: 12pt
mainfont: Times New Roman
mathfont: Times New Roman
---

@@

library(checkpoint)
checkpoint("2018-11-03")

@@

@@

```{r}

@@

First, if 10 out of 10 heads are reported, you should learn two things. On the one hand, there are
probably many liars in your sample. On the other hand, probably a lot of coins really landed heads.
The probability distribution in equation (\ref{eqn:truncated-binom}) does not take account of this.
In particular, for *R = N*, it reduces to the binomial distribution. This is not correct. For
example, in my simulations with $P = 0.5$, the overall probability that there were 10 heads when $R
= N = 10$ was about 1 in `r inv_prob_h`, not 1 in 1024. When $\lambda = 0.2$, it is about 1 in 
`r inv_prob_h_l20`. The same is true at the other end of the scale: when *R* is low, we indeed learn
that $T \le R$ but also that the overall distribution is not shaped like the prior.

Second, more importantly, the GSV approach estimates *Lies*, the distribution of the proportion of 
*lies actually told*, among the subsample of people who saw tails. But we are not usually interested
in the proportion of lies actually told. We care about the probability that people in the
sample would lie if they saw tails --  $\lambda$ in equation (\ref{eqn:lambda}). This $\lambda$ can
be interpreted in different ways. It could be that on seeing a tail, each person in the sample lies
with probability $\lambda$. Or it could be that the sample is drawn from a population of whom
$\lambda$ are (always) liars, and $1 - \lambda$ are truth-tellers.

@@

*Lies* can be treated as an estimate of $\lambda$. It is unbiased, because the
probability that people would lie if they saw tails is independent of the coin flip they actually
see. But it can be a very noisy estimate. 
Again, suppose 10 heads out of 10 are reported, and 9 heads were really observed. *Lies* is 100%. 
But it is 100% of just one person.

@@

The second problem means that the GSV confidence intervals are not correct for $\lambda$. We can
tell that the second problem matters, because the confidence interval coverage of $\lambda$ is much
worse than coverage of *Lies*. The GSV estimator is also a biased estimator of $\lambda$.

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so *T* is high and the true sample size $N - T$ is low. Table \ref{tab:coverage-rn}
shows this. It splits the simulations by the proportion of reported heads, *R/N*. Coverage levels
fall off sharply as *R/N* increases. Note that for fair coin flips, *R/N* is typically greater than
0.5, in the simulations and in reality.

@@

@@

We can also compare the accuracy of point estimates of *Liars*. Table
\ref{tab:point-estimates} shows the mean squared error for methods by different *N*. For low *N*,
the best method is Bayesian and the worst is Frequentist, with GSV in between. When *N* gets large all
methods give about the same estimates and are equally accurate.

@@

sim_est %>% 
      filter(CI == 90) %>% 
      mutate(
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      group_by(Method, N) %>% 
      summarize(mean_sq_error = mean(error_sq, na.rm = TRUE)) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean squared errors by method and $N$") %>% 
      set_label("tab:point-estimates") %>% 
      theme_plain()
```

>>>>

---
title: "True Lies"
subtitle: "Comment on Garbarino, Slonim and Villeval (2018)"
author: "David Hugh-Jones"
date: "`r Sys.Date()`"
header-includes:
  - \usepackage{setspace}
  - \usepackage{placeins}
  - \onehalfspacing
  - \usepackage{fontspec}
output:
  pdf_document:
    md_extensions: +raw_attribute
    latex_engine: xelatex
  tufte::tufte_handout:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_handout2:
    latex_engine: xelatex
    md_extensions: +raw_attribute
  bookdown::tufte_html2: 
    pandoc_args: ['--self-contained']
    keep_intermediates: true
bibliography: bibliography.bib
link-citations: yes
fontsize: 12pt
mainfont: Times New Roman
mathfont: Times New Roman
---

@@

library(checkpoint)
checkpoint("2018-11-03", scanForPackages = cache_results)

@@

# check we aren't too far from their calculations. 
with(sim_est, stopifnot(
        max(abs(EV.recalc - EV.gsv)) < 0.0001,
        mean(abs(LB.recalc - LB.gsv)) < 0.05,
        mean(abs(UB.recalc - UB.gsv)) < 0.05
      ))
```

@@

```{r heads-10-10-calcs}

@@

First, if 10 out of 10 heads are reported, you should learn two things. On the one hand, there are
probably many liars in your sample. On the other hand, probably a lot of coins really landed heads.
The probability distribution in equation (\ref{eqn:truncated-binom}) does not take account of this.
In particular, for *R = N*, it reduces to the binomial distribution. This is not correct. For
example, in my simulations with $P = 0.5$, the overall probability that there were 10 heads when $R
= N = 10$ was about 1 in `r inv_prob_h`, not 1 in 1024. When $\lambda = 0.2$, it was about 1 in 
`r inv_prob_h_l20`. The same is true at the other end of the scale: when *R* is low, we indeed learn
that $T \le R$ but also that the overall distribution is not shaped like the prior.

Second, more importantly, the GSV approach estimates *Lies*, the distribution of the proportion of 
*lies actually told*, among the subsample of people who saw tails. But we are not usually interested
in the proportion of lies actually told. We care about the probability that people in the
sample would lie if they saw tails --  $\lambda$ in equation (\ref{eqn:lambda}). This $\lambda$ can
be interpreted in different ways. It could be that on seeing a tail, each person in the sample lies
with probability $\lambda$. Or it could be that the sample is drawn from a population of whom
$\lambda$ are (always) liars, and $1 - \lambda$ are truth-tellers. *Lies*, on the other hand,
has no interpretation in the population, because the rest of the population has no chance to tell a lie
in the experiment. Indeed, when GSV run Monte Carlo simulations, they themselves check their
confidence intervals against $\lambda$ (which they call $Q$).

@@

*Lies* can be treated as an estimate of $\lambda$. It is unbiased: it is an estimate of 
$\lambda$ in the random sample of $N - T$ people who saw tails. But it can be a very noisy estimate. 
Again, suppose 10 heads out of 10 are reported, and 9 heads were really observed. *Lies* is 100%. 
But it is 100% of just one person.

@@

The second problem means that the GSV confidence intervals are not correct for $\lambda$. We can
tell that the second problem matters, because the confidence interval coverage of $\lambda$ is much
worse than coverage of *Lies*. The GSV estimator is also a biased estimator of $\lambda$, as we will
see.

@@

The problem is especially bad when there are many reports of heads. In this case there were probably
many true heads, so *T* is high and the true sample size $N - T$ is low. Table \ref{tab:coverage-rn}
shows this. It splits the simulations by the proportion of reported heads, *R/N*. Coverage levels
fall off sharply as *R/N* increases. For fair coin flips, *R/N* is typically greater than
0.5, in the simulations and in reality.

@@

```{r benndorf-recalc}
bennd <- update_prior(57, 98, 5/6, dunif)
bennd_bounds <- dist_hdr(bennd, 0.95)
bennd_bounds <- pct_format(bennd_bounds)
bennd_trad <- binom.test(57, 98)$conf.int
bennd_trad <- pct_format(bennd_trad)
```

@benndorf2017experienced use the GSV method to calculate confidence intervals for the proportion of
liars in a lying task with a die roll ($P = 5/6$). From 57 reports of the best outcome, out of 98
subjects, they calculate a lying rate of 49.68%, with a 95% CI of (45.3%, 53.95%). Using the
Bayesian method with a uniform prior, the confidence interval becomes (`r bennd_bounds[1]`, 
`r bennd_bounds[2]`), about twice as big. Obviously, this matters for e.g. sample size calculations.

@@

We can also compare the accuracy of point estimates of *Liars*. Table \ref{tab:bias} shows bias
(the estimated value minus the true value of $\lambda$) for different methods by different *N*.
The Bayesian method is always the least biased until $N = 500$, and the GSV method is the most biased.

Table \ref{tab:mse} shows the mean squared error for methods by different *N*. For low *N*,
the best method is Bayesian and the worst is Frequentist, with GSV in between. When *N* gets large all
methods give about the same estimates and are equally accurate.

@@

bias_mse <- sim_est %>% 
      filter(CI == 90) %>% 
      mutate(
        Method = recode(method, gsv = "GSV", simple = "Frequentist", ubayes = "Bayesian"),
        method = NULL
      ) %>% 
      group_by(Method, N) %>% 
      summarize(
        mean_sq_error = mean(error_sq, na.rm = TRUE),
        bias          = mean(error, na.rm = TRUE)
      )

bias_mse %>% 
      select(-mean_sq_error) %>% 
      spread(N, bias, sep = ": ") %>% 
      ungroup() %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean bias by method and $N$") %>% 
      set_label("tab:bias") %>% 
      theme_plain()

bias_mse %>% 
      select(-bias) %>% 
      spread(N, mean_sq_error, sep = ": ") %>% 
      ungroup() %>% 
      as_huxtable(add_colnames = TRUE) %>% 
      set_caption("Mean squared errors by method and $N$") %>% 
      set_label("tab:mse") %>% 
      theme_plain()

# Power tests

```{r power-tests}
power_l_25 <- power_calc(N = 100, P = 0.5, lambda = 0.25)
power_l_10 <- power_calc(N = 100, P = 0.5, lambda = 0.10)
```

GSV argue that existing sample sizes may be too small to reject "no lying" ($\lambda = 0). 
We can check this with the Bayesian method by simulations. With a uniform prior and an $N$ of 100, 
the Bayesian method has `r pct_format(power_l_25)` power to detect $\lambda$ of 25% and 
`r pct_format(power_l_10)` power to detect $\lambda$ of 10%. So, on this we agree.
